Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-12 paper/paper.tex
adapted kw
added submitted rev

% official NIPS keywords
\textbf{\\keywords}: systems neuroscience, artificial intelligence, reinforcement learning, mind-wandering

@@

In short, predictive coding processes
conceptualize updates of the internal representation of the environment
to best accommodate and prepare the organism
for processing the constant influx of sensory stimuli and
performing action on the environment.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account of DMN function based on MDPs.
Importantly,
a generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both functional accounts are supported by
neuroscientific evidence that suggest
the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
with the biological purpose to
generalize from the past to new experiences.
Neuroanatomically, axonal
back projections indeed outnumber by far the axonal connections mediating feedforward input processing
in the monkey brain and probably also in humans
\citep{salin1995corticocortical}.
These many and diverse top-down modulations
from higher onto downstream cortical areas
can inject prior knowledge
at every stage of processing environmental information.
%
Moreover,
both accounts provide a parsimonious explanation for why the
human brain's processing load devoted to incoming information decreases
when the environment becomes predictable.
This is because the internal generative
model only requires updates after discrepancies have occurred between
environmental reality and its internally reinstantiated representation.
Increased computation resources are however allocated
when unknown stimuli or
unexpected events are encountered by the organism.
The predictive coding and MDP account hence
naturally evoke a mechanism of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unanticipated environmental challenges.

@@

\subsection{The semantic account}
This frequently proposed cognitive account to explain DMN function revolves
around forming logical associations and abstract analogies between
experiences and conceptual knowledge derived from past behavior
\citep{bar2007proactive, binder1999conceptual, constantinescu2016organizing}.
Analogies might naturally tie incoming new sensory stimuli to
explicit world knowledge (i.e., semantics)
\citep{bar2009proactive}.
The encoding of complex environmental features could thus be facilitated
by association to known similar states.
%
Going beyond isolated meaning and concepts extracted from the world,
semantic building blocks may need to get recombined to enable
mental imagery to (fore)see never-experienced scenarios.
As such, semantic knowledge would be an important ingredient for optimizing behavior
by constantly simulating possible future scenarios
\citep{boyer2008evolutionary, binder2011neurobiology}.
Such cognitive processes can afford
the internal construction and elaboration of necessary information
that is not presented in the immediate sensory environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imagining the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.

@@

\subsection{The sentinel account}
Regions of the DMN have been proposed to process the
experienced or expected relevance of environment cues
\citep{montague2006imaging}.
Processing self-relevant information was perhaps the first
functional account that was proposed for the DMN
\citep{gusnard2001medial, raichle2001pnas}.
Since then,
many investigators have speculated that neural activity in the DMN
may reflect the brain's continuous tracking of
relevance in the environment, such as spotting predators,
as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
According to this cognitive account, the human brain's baseline maintains
a ``radar'' function to
detect subjectively relevant cues and unexpected events in the environment.
Propositions of a sentinel function to underlie DMN activity
have however seldom detailed
the mechanisms of
how attention and memory resources are exactly reallocated when
encountering a self-relevant environmental stimulus.
Instead,
in the present MDP account,
promising action trajectories
are recursively explored by the human DMN. Conversely,
certain branches of candidate action trajectories
are detected to be less worthy to get explored.
This mechanism, expressed by the Bellman equation,
directly implies stratified allocation of attention and working memory load
over relevant cues and events in the environment.

@@

In particular,
environmental cues that are especially important for humans are frequently of
social nature. This may not be surprising
given that
the complexity of the social systems
is likely to be a human-defining property
\citep{tomasello2009cultural, dunbar2007science}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
In fact, social topics probably amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}.
Mind-wandering at daytime and dreams during sleep
are also rich in stories about people and
the complex interactions between them.
%
In line with this, DMN activity was advocated to be specialized in
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default},
cats \citep{popa2009contrasting},
and rats \citep{lu2012rat}, three species with
social capacities that can be expected to be less advanced than in humans.

>>>>

% official NIPS keywords
\textbf{\\keywords}: systems neuroscience, artificial intelligence, mind-wandering

@@

In short, predictive coding processes
conceptualize updates of the internal representation of the environment
to best accommodate and prepare the organism
for processing the constant influx of sensory stimuli and
performing action on the environment (Fig. \ref{fig:VennDiagram}).
There are hence a number of common properties
between the predictive coding account
and the proposed formal account of DMN function based on MDPs.
Importantly,
a generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both functional accounts are supported by
neuroscientific evidence that suggest
the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
with the biological purpose to
generalize from the past to new experiences.
Neuroanatomically, axonal
back projections indeed outnumber by far the axonal connections mediating feedforward input processing
in the monkey brain and probably also in humans
\citep{salin1995corticocortical}.
These many and diverse top-down modulations
from higher onto downstream cortical areas
can inject prior knowledge
at every stage of processing environmental information.
%
Moreover,
both accounts provide a parsimonious explanation for why the
human brain's processing load devoted to incoming information decreases
when the environment becomes predictable.
This is because the internal generative
model only requires updates after discrepancies have occurred between
environmental reality and its internally reinstantiated representation.
Increased computation resources are however allocated
when unknown stimuli or
unexpected events are encountered by the organism.
The predictive coding and MDP account hence
naturally evoke a mechanism of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unanticipated environmental challenges.

@@

\subsection{The semantic account}
This frequently proposed cognitive account to explain DMN function revolves
around forming logical associations and abstract analogies between
experiences and conceptual knowledge derived from past behavior
\citep{bar2007proactive, binder1999conceptual, constantinescu2016organizing}.
Analogies might naturally tie incoming new sensory stimuli to
explicit world knowledge (i.e., semantics, Fig. \ref{fig:VennDiagram})
\citep{bar2009proactive}.
The encoding of complex environmental features could thus be facilitated
by association to known similar states.
%
Going beyond isolated meaning and concepts extracted from the world,
semantic building blocks may need to get recombined to enable
mental imagery to (fore)see never-experienced scenarios.
As such, semantic knowledge would be an important ingredient for optimizing behavior
by constantly simulating possible future scenarios
\citep{boyer2008evolutionary, binder2011neurobiology}.
Such cognitive processes can afford
the internal construction and elaboration of necessary information
that is not presented in the immediate sensory environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imagining the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.

@@

\subsection{The sentinel account}
Regions of the DMN have been proposed to process the
experienced or expected relevance of environment cues
\citep{montague2006imaging}.
Processing self-relevant information was perhaps the first
functional account that was proposed for the DMN
\citep{gusnard2001medial, raichle2001pnas}.
Since then,
many investigators have speculated that neural activity in the DMN
may reflect the brain's continuous tracking of
relevance in the environment, such as spotting predators,
as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
According to this cognitive account, the human brain's baseline maintains
a ``radar'' function to
detect subjectively relevant cues and unexpected events in the environment
(Fig. \ref{fig:VennDiagram}).
Propositions of a sentinel function to underlie DMN activity
have however seldom detailed
the mechanisms of
how attention and memory resources are exactly reallocated when
encountering a self-relevant environmental stimulus.
Instead,
in the present MDP account,
promising action trajectories
are recursively explored by the human DMN. Conversely,
certain branches of candidate action trajectories
are detected to be less worthy to get explored.
This mechanism, expressed by the Bellman equation,
directly implies stratified allocation of attention and working memory load
over relevant cues and events in the environment.

@@

In particular,
environmental cues that are especially important for humans are frequently of
social nature. This may not be surprising
given that
the complexity of the social systems
is likely to be a human-defining property
\citep{tomasello2009cultural, dunbar2007science}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
In fact, social topics probably amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}.
Mind-wandering at daytime and dreams during sleep
are also rich in stories about people and
the complex interactions between them.
%
In line with this, DMN activity was advocated to be specialized in
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default},
cats \citep{popa2009contrasting},
and rats \citep{lu2012rat}, three species with
social capacities that can be expected to be less advanced than in humans
\citep{mars2012_social_brain}.
